---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```



```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(formattable)
library(ggrepel)
library(corrplot)
library(Hmisc)
library(tidyverse)
library(googlesheets4)
library(rlang)
library(ggplot2)
library(data.table)
library(DT)
library(knitr)
library(kableExtra)
library(ggrepel)
library(plotly)

#Here are the links to data sources, in order as they appear in the "Cost of Doing Business - Full Data" database. Apologies for the disorganizaton, I've just been putting this together on the fly

#Urbanization (Column 2) - https://fivethirtyeight.com/features/how-urban-or-rural-is-your-state-and-what-does-that-mean-for-the-2020-election/

#Tax Burden (Columns 3- 6) - https://wallethub.com/edu/states-with-highest-lowest-tax-burden/20494

#Tax Ranking by State (Columns 7 - 12) - https://taxfoundation.org/2021-state-business-tax-climate-index/

#Columns 13-32 are the Highway Report Data 

#CNBC (Columns 33 - 42) - https://www.cnbc.com/2021/07/13/americas-top-states-for-business.html

#RPP (Column 43 )- https://apps.bea.gov/iTable/iTable.cfm?reqid=70&step=1&acrdn=8



```

## Why are some states doing so poorly on the AHR? 
This analysis seeks to explore the relationship between the finance score and the costs associated with transportation construction and maintenance in urban vs rural states 

```{r}

historical_data <- readr::read_csv("https://raw.githubusercontent.com/spurnell/AHR_Density_Report/main/R_Feeder_%20MAIN_Historical%20-%20Main%20Database.csv")


historical_data <- historical_data %>% 
  filter(!Year == 2017) %>% 
  filter(!Year == 2010)

historical_data <- historical_data[,c(1:2, 5:9)]

historical_data <- as.data.frame(lapply(historical_data, function(y) gsub(",", "", y)))

historical_data <- rename(historical_data, State = . )

historical_data[2:7] <- as.numeric(unlist(historical_data[2:7]))

per_miles <- function(x) {
  x/historical_data$State.Controlled.Lane.Miles
  
}
  
state_controlled_LM <- historical_data %>% 
  mutate(across(c(4,5,6,7), per_miles))

state_controlled_LM <- state_controlled_LM[4:7]

state_controlled_LM <- state_controlled_LM %>% 
  rename( Capital.Disbursements.Per.LM = Capital.Disbursements,
          Maintenence.Disbursements.Per.LM = Maintenence.Disbursements,
          Administrative.Disbursements.Per.LM = Administrative.Disbursements,
          Total.Disbursements.Per.LM = Total.Disbursements
          )

historical_data <- cbind(historical_data, state_controlled_LM)

```


## Sum of Spending in All States
```{r}
#Calculating spending sums by year
raw_spending_by_year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    dplyr::summarize(Total_Spending = (sum(Total.Disbursements)),
                     Admin_Spending = sum(Administrative.Disbursements),
                     Maint_Spending = sum(Maintenence.Disbursements),
                     Capital_Spending = sum(Capital.Disbursements)) %>% 
    mutate(Total_Spending = formattable::currency(Total_Spending))  %>% 
    mutate(Admin_Spending = formattable::currency(Admin_Spending)) %>% 
    mutate(Maint_Spending = formattable::currency(Maint_Spending)) %>% 
    mutate(Capital_Spending = formattable::currency(Capital_Spending))
             
  
}
              
Spending_by_year <- historical_data %>% 
  raw_spending_by_year()


#plotting spending
ggplot2::ggplot(data = Spending_by_year, aes(Year)) +
  geom_line(aes( y = Total_Spending, colour = "Total")) +
  ylab("Total Spending") +
  scale_y_continuous(labels = comma)
    

ggplot2::ggplot(data = Spending_by_year, aes(Year)) +
  geom_line(aes(y = Admin_Spending, colour = "Admin")) +
  geom_line(aes(y = Maint_Spending, colour = "Maint")) +
  geom_line(aes(y = Capital_Spending, colour = "Capital")) +
  ggplot2::ylab("Other Spending") +
  scale_y_continuous(labels = comma)

  


#percentage Increases
percent_increase <- function(x) {
  
  Total_percent_Increase <- (((Spending_by_year[13,x]) - (Spending_by_year[1,x])) / Spending_by_year[1,x]) 
  
  Total_percent_Increase <- formattable::percent(Total_percent_Increase)
  
}

Total_increase <- percent_increase(2)
Admin_increase <- percent_increase(3)
Maint_increase <- percent_increase(4)
Capit_increase <- percent_increase(5)

Increases <- cbind(Total_increase, Admin_increase, Maint_increase, Capit_increase)

Increases <- formattable::percent(Increases)

#Tables for total sum spending
kableExtra::kbl(Spending_by_year) %>% 
  kable_styling()

print(Increases)

#calculating per mile averages 

permile_avg_by_year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    dplyr::summarize(Mean_Total_Spending = mean(Total.Disbursements.Per.LM),
                     Mean_Admin_Spending = mean(Administrative.Disbursements.Per.LM),
                     Mean_Maint_Spending = mean(Maintenence.Disbursements.Per.LM),
                     Mean_Capital_Spending = mean(Capital.Disbursements.Per.LM))
  
}

Spending_permile_avg <- historical_data %>% 
  permile_avg_by_year()



#Table for Spending per Lane Mile
kableExtra::kbl(Spending_permile_avg, caption = "Spending (in Millions) per Lane Mile") %>% 
  kable_styling()

#plotting per mile 
ggplot2::ggplot(data = Spending_permile_avg, aes(Year)) +
  geom_line(aes(y = Mean_Total_Spending, colour = "Total")) +
  geom_line(aes(y =  Mean_Admin_Spending, colour = "Admin")) +
  geom_line(aes(y =  Mean_Maint_Spending, colour = "Maint")) +
  geom_line(aes(y = Mean_Maint_Spending, colour = "Capital")) +
  ylab("Mean Spending per Mile (In Millions)")



#calculating variance of disbursements per lane mile 
variance_by_year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    dplyr::summarize(Total_Variance = var(Total.Disbursements.Per.LM),
                     Admin_Variance = var(Administrative.Disbursements.Per.LM),
                     Maint_Variance = var(Maintenence.Disbursements.Per.LM),
                     Capit_Variance = var(Capital.Disbursements.Per.LM))
  
}

yearly_variance <- historical_data %>% 
  variance_by_year() 

kableExtra::kbl(yearly_variance) %>% 
  kable_styling()


#plotting variance of disbursements per lane miles
ggplot(data = yearly_variance, aes(Year)) +
  geom_line(aes(y = Total_Variance, colour = "Total")) 


ggplot(data = yearly_variance, aes(Year)) +
  geom_line(aes(y = Admin_Variance, colour = "Admin")) +
  geom_line(aes(y = Maint_Variance, colour = "Maint")) +
  geom_line(aes(y = Capit_Variance, colour = "Capital"))




#calculating per mile maxes
max_by_year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    dplyr::summarize(Total_Max = max(Total.Disbursements.Per.LM),
                     Admin_Max = max(Administrative.Disbursements.Per.LM),
                     Maint_Max= max(Maintenence.Disbursements.Per.LM),
                     Capit_Max = max(Capital.Disbursements.Per.LM))
  
}

yearly_max <- historical_data %>% 
  max_by_year()

#plotting per mile maxes 
ggplot(yearly_max, aes(Year)) +
  geom_line(aes( y = Total_Max, colour = "Total"))

ggplot(yearly_max, aes(Year)) + 
  geom_line(aes(y = Admin_Max, colour = "Admin")) +
  geom_line(aes(y = Maint_Max, colour = "Maint")) +
  geom_line(aes(y = Capit_Max, colour = "Capital"))



#Calculating Finance Scores 
get_scores <- function(.data, sum_col, old_col, new_col) {
  
  old_col <- enquo(old_col)
  new_col <- enquo(new_col)
  new_col_name <- quo_name(new_col)
  sum_col <- enquo(sum_col)
  
  
historical_data <- historical_data %>% 
  group_by(Year) %>% 
  mutate(!!new_col_name := !!old_col/
           (sum(!!sum_col) / sum(State.Controlled.Lane.Miles)))

}

historical_data <- historical_data %>% 
  get_scores(Capital.Disbursements, Capital.Disbursements.Per.LM, Capital.Score)

historical_data <- historical_data %>% 
  get_scores(Maintenence.Disbursements, Maintenence.Disbursements.Per.LM, Maintenence.Score) 

historical_data <-  historical_data %>% 
  get_scores(Administrative.Disbursements, Administrative.Disbursements.Per.LM, Administrative.Score)

historical_data <- historical_data %>% 
  get_scores(Total.Disbursements, Total.Disbursements.Per.LM, Total.Score)

historical_data <- historical_data %>% 
  rowwise() %>% 
  ungroup() %>% 
  mutate(Finance_Score = Capital.Score +
                          Maintenence.Score +
                          Administrative.Score +
                          Total.Score)


historical_data$Finance_Score <- historical_data$Finance_Score/4


#Calculating Finance Score Variance 
Finance_variance_by_year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    dplyr::summarize(Total_Variance = var(Total.Score),
                     Admin_Variance = var(Administrative.Score),
                     Maint_Variance = var(Maintenence.Score),
                     Capit_Variance = var(Capital.Score))
  
}

Finance_Score_Variance <- historical_data %>% 
  Finance_variance_by_year()



print(Finance_Score_Variance)

#plotting finance score variance 
ggplot2::ggplot(data = Finance_Score_Variance, aes(Year)) +
  geom_line(aes( y = Total_Variance, colour = "Total")) +
  geom_line(aes(y = Admin_Variance, colour = "Admin")) +
  geom_line(aes(y = Maint_Variance, colour = "Maint")) +
  geom_line(aes(y = Capit_Variance, colour = "Capital"))
  
  

#calculating finance maxes
Finace_Max_by_Year <- function(.data) {
  
  .data %>% 
    dplyr::group_by(Year) %>% 
    summarize(Total_max = max(Total.Score),
              Admin_max = max(Administrative.Score),
              Maint_max = max(Maintenence.Score),
              Capit_max = max(Capital.Score))
  
}


Finance_Maxes <- historical_data %>% 
  Finace_Max_by_Year()
  
#plotting finance maxes 
ggplot2::ggplot(Finance_Maxes, aes(Year)) +
  geom_line(aes( y = Total_max, colour = "Total")) +
  geom_line(aes(y = Admin_max, colour = "Admin")) +
  geom_line(aes(y = Maint_max, colour = "Maint")) +
  geom_line(aes(y = Capit_max, colour = "Capital"))
  

```


## Greater Variance In Finance Scores than Performance Scores

New Jersey has by far the worst Finance Score, almost double

```{r}
Full_Data <- readr::read_csv("https://raw.githubusercontent.com/spurnell/AHR_Density_Report/main/Cost%20of%20doing%20business%20-%20Full%20Data.csv")

Full_Data <- as.data.frame(Full_Data)

variance <- Full_Data[, c(1,13,15)]

print(variance[order(variance[,2], decreasing = T),  ])

```

Finance STDV, Variance

```{r}

finance_score_dev <- sd(variance[,2])

print(finance_score_dev)

finance_score_var <- var(variance[,2])

print(finance_score_var)


```
Performance STDV, Variance

```{r}
perforance_score_dev <- sd(variance[,3])

print(perforance_score_dev)

perforance_score_var <- var(variance[,3])

print(perforance_score_var)

```



## Tax, Finance, and Performance Ranking Matrix

Shows the correlations between all rankings  Notice that Finance Rank correlates with both cost of living and doing business but nothing else. Neither is correlated with performance rank either.  

```{r pressure, echo=FALSE}

matrix_plot <- readr::read_csv("https://raw.githubusercontent.com/spurnell/AHR_Density_Report/main/Cost%20of%20doing%20business%20-%20matrix.csv")

matrix_plot <- as.data.frame(matrix_plot)

matrix_plot2 <- matrix_plot[,c(1:4,19,28)]

res <- cor(matrix_plot2)

corrplot(res)
```

## Multiple Regression Analysis

Explores statistical relationships between Finance, Performance, and Localized Costs Ranks

Here's the CNBC Definitions:

  - **Cost of Doing Business:**  We measure the strength of each state’s business tax climate, as well as tax burdens for various types of businesses and facilities. We also measure wage and utility costs, as well as the cost of office and industrial space. And we consider incentives and tax breaks that states offer to reduce business costs, with special emphasis on incentives targeted toward development in disadvantaged communities.
  
  - **Education:**  We consider the number of higher education institutions in each state as well as long-term trends in state support for higher education. We also consider historically Black colleges and universities (HBCUs), which companies are increasingly seeking to partner with. We look at multiple measures of K-12 education including test scores, class size and spending. We also look at life-long learning opportunities in each state.
  
  - **Workforce:** We measure the educational attainment of each state’s working-age population, as well as which states are attracting college-educated workers and which states are losing them. With skilled workers in particular demand, we consider each state’s concentration of science, technology, engineering and math (STEM) workers. We measure worker productivity based on economic output per job. We look at union membership and right to work laws. And we measure the availability of workers, as well as the diversity of each state’s workforce.
  
  - **Cost of Living:** The cost of living helps drive the cost of doing business. From housing to food and energy, wages go further when the cost of living is low. We measure the states based on an index of costs for basic items.

## Overall Rank


```{r, anlysis}

cost_exploration <- readr::read_csv("https://raw.githubusercontent.com/spurnell/AHR_Density_Report/main/Cost%20of%20doing%20business%20-%20Full%20Data.csv", name_repair = "unique")

cost_exploration <- as.data.frame(lapply(cost_exploration, function(y) gsub(",", "", y)))

cost_exploration[,2:43] <- as.numeric(unlist(cost_exploration[,2:43]))

linear_model <- with(cost_exploration, lm(Overall.Rank ~ EDUCATION +
                                                          WORKFORCE +
                                                          Cost.of.Doing.Business +
                                                          Cost.of.Living))




summary(linear_model)





```

## Finance Rank 

```{r}

linear_model2 <- with(cost_exploration, lm(Finance.Rank ~ EDUCATION +
                                                          WORKFORCE +
                                                          Cost.of.Doing.Business +
                                                          Cost.of.Living))
summary(linear_model2)

```

## RPP + Overall Score 

```{r}

linear_model3 <- with(cost_exploration, lm(Overall.Score ~ RPP ))

summary(linear_model3)

plot(cost_exploration$Overall.Score, cost_exploration$RPP)


```

##  ******* Finance Score + RPP ************

```{r}

linear_model4 <- with(cost_exploration, lm(Finance.Score ~ RPP))

summary(linear_model4)

plot(cost_exploration$Finance.Score, cost_exploration$RPP)

```


# Finance Score + Performance + RPP

```{r}

linear_model5 <- with(cost_exploration, lm(Finance.Score ~ Performance.Score + RPP))

summary(linear_model5)

```

## Cost of Living vs Tax Burden 

```{r, warning=F}
cost_of_living <- readr::read_csv("https://raw.githubusercontent.com/spurnell/AHR_Density_Report/main/advisorsmith_cost_of_living_index%20-%20advisorsmith_cost_of_living_index.csv")

cost_of_living <- as.data.frame(lapply(cost_of_living, function(y) gsub(",", "", y)))

cost_of_living <- cost_of_living %>% 
  filter(!State == "DC")

cost_of_living$State <- state.name[match(cost_of_living$State, state.abb)]

cost_of_living <- cost_of_living %>% 
  group_by(State) %>% 
  summarise(avg_cost_of_living = round(mean(Cost.of.Living.Index), 2))


Full_Data_smithadv <- merge(Full_Data, cost_of_living, by = "State")

ggplot(Full_Data_smithadv, aes(x=Urbanization, y =avg_cost_of_living)) +
  geom_point() +
  ggrepel::geom_text_repel(
    label = Full_Data_smithadv$State
  ) 

state_names <- Full_Data_smithadv$State
  
fig <- plotly::plot_ly(Full_Data_smithadv, 
                       x=~Urbanization, y =~avg_cost_of_living,
                       mode = 'text',
                       text = ~Full_Data_smithadv$State)

fig


fig2 <- plotly::plot_ly(Full_Data_smithadv, 
                       x=~Urbanization, y =~`Finance Score`,
                       mode = 'text',
                       text = ~Full_Data_smithadv$State)

fig2


```




## Results
 
The above results show that when controlling for differences in Workforce and Education, the cost of doing business has a significant impact on the Overall Score, which is the best indicator 